---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nert

<!-- badges: start -->
[![R-CMD-check](https://github.com/AAGI-AUS/nert/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/AAGI-AUS/nert/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/AAGI-AUS/nert/graph/badge.svg)](https://app.codecov.io/gh/AAGI-AUS/nert)
<!-- badges: end -->

The goal of {nert} is to provide access to Australian TERN (Terrestrial Ecosystem Research Network) data in your R session.

## Installation

You can install the development version of {nert} from [GitHub](https://github.com/AAGI-AUS/nert) with:

```{r install, eval=FALSE}
if (!require("pak")) {
  install.packages("pak")
}

pak::pak("AAGI-AUS/nert")
```

## Example: reading a COG as a spatial object

This is a basic example which shows you how you can fetch one day's data from the SMIPS data (currently the only supported data set in TERN) and visualise it:

```{r example_cog}
library(nert)
r <- read_smips(day = "2024-01-01")

plot(r)
```

## Extract Values Given Lat/Lon Values

Extract Soil Moisture for Corrigin and Merriden, WA and Tamworth, NSW given latitude and longitude values for each.

```{r}
library(terra)
df <- structure(
  list(
    location = c("Corrigin", "Merredin", "Tamworth"),
    x = c(117.87, 118.28, 150.84),
    y = c(-32.33, -31.48, -31.07)
  ),
  row.names = c(NA, -3L),
  class = "data.frame"
)

cog_df <- extract(x = r, y = df[, c("x", "y")], xy = TRUE)

cog_df <- cbind(df$location, cog_df)
names(cog_df) <- c("location", "ID", "smips_totalbucket_mm_20240101", "x", "y")
cog_df
```

## Citing {nert}

To cite nert:

```{r cite}
citation("nert")
```
